---
title: "Question 3"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
rm(list=ls())
library(dplyr)
library(tools)
library(ggplot2)
library(fixest)
library(sandwich)
library(pander)
library(kableExtra)
```


```{r}
# Prepare the data
df <- read.table("PS2Data/CARD.raw", 
                 quote="\"", comment.char="")

name_string <- "id   nearc2    nearc4    educ      age       fatheduc  motheduc  
weight   momdad14  sinmom14  step14    reg661    reg662    reg663    reg664   
reg665 reg666    reg667    reg668    reg669    south66   black     smsa   south    
smsa66    wage      enroll    KWW       IQ        married   libcrd14  exper    
lwage     expersq   "

name_string <- gsub("[\r\n]", " ", name_string)
name_string <- strsplit(name_string, " ")

name_vec <- vector()
for (i in name_string[[1]]) {
  if (i != "1" & i != ""){
    name_vec <- append(name_vec, i)
  }
}
colnames(df) <- name_vec
```

# a

In the table below, the $iid$ and $robust$ standard errors are quite similar ($robust$ standard errors are usually a little larmer), so the inference (ttest and pvalue) is also quite similar.

The $iid$ standard error assumes that the variance of the error term is constant and does not depend on independent variables. However, the $robust$ standard error does not assume this, so it can work under both homoskedasticity and heteroskedasticity.
```{r, results='hide'}
# Homoskedastic
homo <- feols(lwage ~ educ + exper + expersq + black + south + smsa + smsa66 +
              reg661 + reg662 + reg663 + reg664 + reg665 + reg666 + reg667 + reg668,
              se = "iid", df)

# Heteroskedastic
hetero <- feols(lwage ~ educ + exper + expersq + black + south + smsa + smsa66 +
                reg661 + reg662 + reg663 + reg664 + reg665 + reg666 + reg667 + reg668,
                se = "hc1", df)

out <- etable(homo, hetero, tex = TRUE, se.row = TRUE) 
```

```{r, include=knitr::is_latex_output()}
knitr::asis_output(c("\\begin{center}", out, "\\end{center}")) 
```

\newpage
# b

There exists a practically and statistically significant partial correlation between $educ$ and $nearc4$: the coefficient of $nearc4$ is $0.32$, which means indivisuals near 4 yr college have additional $0.32$ years of schooling. It's significant under $1\%$ level with both $iid$ and $robust$ standard errors.

For some variables, the $robust$ standard error is a little larger, and for other variables, the $robust$ standard error is a little smaller. But they are still quite similar and the inference (ttest and pvalue) gives same conclusions. For $near4$, the $robust$ standard error is a little smaller, but the coefficient is both significant under $1\%$ level.
```{r, results='hide'}
reduce_homo <- feols(educ ~ nearc4 + exper + expersq + black + south + smsa + smsa66 +
                     reg661 + reg662 + reg663 + reg664 + reg665 + reg666 + reg667 + reg668,
                     se = "iid", df)

reduce_hetero <- feols(educ ~ nearc4 + exper + expersq + black + south + smsa + smsa66 +
                      reg661 + reg662 + reg663 + reg664 + reg665 + reg666 + reg667 + reg668,
                      se = "hc1", df)

out <- etable(reduce_homo, reduce_hetero, tex = TRUE, se.row = TRUE) 
```

```{r, include=knitr::is_latex_output()}
knitr::asis_output(c("\\begin{center}", out, "\\end{center}")) 
```

\newpage
# c

The IV estimate has wider CI and the lower bound is closer to 0, hence it's more conservative. However, the estimates from both IV and OLS are significant on $95 \%$ level since the lower bounds are both larger than 0.
```{r, results='hide'}
iv_c <- feols(lwage ~ exper + expersq + black + south + smsa + smsa66 +
              reg661 + reg662 + reg663 + reg664 + reg665 + reg666 + reg667 + reg668 |
              educ ~ nearc4,
              se = "hc1", df)

out <- etable(iv_c, hetero, tex = TRUE, 
              coefstat = c("confint"),
              headers=list("IV" = 1, "OLS" = 1)) 
```

```{r, include=knitr::is_latex_output()}
knitr::asis_output(c("\\begin{center}", out, "\\end{center}")) 
```

\newpage
# d

The table below presents estimation results from the reduced form. After adding $nearc2$, the coefficient of $nearc4$ is even a littler larger and the $se$ is essentially the same. However, the coefficient of $nearc2$ is way smaller than $nearc4$ and not significant. Therefore, $nearc4$ is more strongly related to $educ$ than $nearc2$. After adding $nearc2$, the adjusted R square also increased a little - the independent variables can jointly explain more variation of $educ$.
```{r, results='hide'}
reduce_d <- feols(educ ~ nearc2 + nearc4 + exper + expersq + black + south + smsa + smsa66 +
                         reg661 + reg662 + reg663 + reg664 + reg665 + reg666 + reg667 + reg668,
                         se = "hc1", df)

out <- etable(reduce_d, reduce_hetero, tex = TRUE, se.row = TRUE) 
```

```{r, include=knitr::is_latex_output()}
knitr::asis_output(c("\\begin{center}", out, "\\end{center}")) 
```

\newpage
After using two IV, the coefficient increases to $0.16$, larger than former IV result ($0.13$) and the se is even smaller, so it does indicates a stronger relationship. The estimate is larger and we are more confident that it's significantly different from 0.

```{r, results='hide'}
iv_d <- feols(lwage ~ exper + expersq + black + south + smsa + smsa66 +
              reg661 + reg662 + reg663 + reg664 + reg665 + reg666 + reg667 + reg668 |
              educ ~ nearc2 + nearc4,
              se = "hc1", df)

out <- etable(iv_d, iv_c, tex = TRUE, 
              headers=list("#IV:2" = 1, "#IV:1" = 1)) 
```

```{r, include=knitr::is_latex_output()}
knitr::asis_output(c("\\begin{center}", out, "\\end{center}")) 
```

\newpage
# e

$IQ$ is significantly correlated with $nearc4$. Intuitively, $IQ$ is also correlated with $education$, so $near4$ might not be a valid IV for $education$ since it does not satisfy the exclusion assumption. The previous IV estimations might be biased.

```{r, results='hide'}
df$IQ <- as.numeric(df$IQ)
ols_e <- feols(IQ ~ nearc4, se = "hc1", df)
out <- etable(ols_e, tex = TRUE) 
```

```{r, include=knitr::is_latex_output()}
knitr::asis_output(c("\\begin{center}", out, "\\end{center}")) 
```

\newpage
# f

After controlling for $region$ variables, the coefficient of $nearc4$ is not significant anymore. Therefore, after controlling for these $region$ variables, $nearc4$ can serve as a valid IV for $education$ since we can exclude the bias led by $IQ$.

```{r, results='hide'}
df$IQ <- as.numeric(df$IQ)
ols_f <- feols(IQ ~ nearc4 + smsa66 + reg661 + reg662 + reg669, se = "hc1", df)
out <- etable(ols_f, ols_e, tex = TRUE) 
```

```{r, include=knitr::is_latex_output()}
knitr::asis_output(c("\\begin{center}", out, "\\end{center}")) 
```

